Predicting the effect of candidate interventions for mitigating disease spread in asymmetric-mixing metapopulations

### 2024-07-25 ms.tex
harmonized figures with analysis variables; added analysis variable table

We evaluate hypothetical interventions to limit the spread of
infection in the Reserve by simulating these interventions using the formal
system explained above. Our main outcome variable is the reduction in simulated
COVID-19 incidence among Indigenous peoples compared to the no-intervention
case, defined 
\begin{equation}
  \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
  \label{eq:DeltaI}
\end{equation}
\noindent
A preliminary step is to find parameterizations that reproduce the observed
outbreaks of B.1.1 and Zeta within the Reserve to obtain a baseline estimate of
infection size $I(\text{no intervention})$. 
To this end we hypothesize that re-infections could have driven the observed
greater spread of the Zeta strain, though it could also be due to the holiday
season around the end of 2020 and beginning of 2021.

Interventions center around reducing the mobility of Indigenous people between
Dourados city and Reserve; the total change in mobility, denoted $\Delta
\gamma$, is the total reduction of the parameter $\gamma_g$ across all 
groups, i.e., 
\begin{equation}
  \Delta \gamma = \sum_g \Delta \gamma_g.
  \label{eq:DeltaGamma}
\end{equation}
\noindent
Different interventions allocate $\Delta \gamma$ (Equation~\ref{eq:DeltaGamma})
in different ways to hypothetically result in different case reductions, $\Delta
I$ (Equation~\ref{eq:DeltaI}). 
One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
equal. Another named strategy is the \emph{80-20} strategy, where the
reduction in mobility in group $g$ is 80\% of the total, and reduction in
group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$. To create alternative
strategies, both the total reduction and the ratio of group-level reduction to 
total reduction can be changed. In addition to total case reduction, we 
measure relative group-level case reduction, $\Delta I_g$, 
for each considered strategy to understand whether case reduction is equitable
between villages and ethnicities under different strategies. 
We finally evaluate the hypothesis that interventions that may be effective for
strains with no re-infection (like B.1.1) will not be equally effective for
strains with re-infection (like Zeta), and vice-versa. 
% This will highlight the importance of accounting for social and
% epidemiological context, alike, in predicting the effectiveness of mitigation
% strategies for pandemic preparedness.

@@

\vspace{0.5em}
\begin{center}
{[ FIGURE \ref{fig:GroupSpecificTravelReduction} ABOUT HERE ]} \\
\end{center}

\vspace{0.5em}
\begin{center}
{[ FIGURE \ref{fig:GlobalTravelReduction} ABOUT HERE ]} \\
\end{center}

@@

@@

@@

\caption{Global reduction in travel to and from the city could have
Infection reduction ($\Delta I$, y-axis) for different groups could be
different as global reduction in travel to and from the city decreases
($\Delta \gamma$, x-axis).}
\label{fig:GlobalTravelReduction}
\vspace{0.3em}
{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ReductionByGroup.pdf}

@@

\caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
\label{fig:GroupSpecificTravelReduction}
\vspace{0.3em}
{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}

>>>>

We evaluate hypothetical interventions to limit the spread of
infection in the Reserve by simulating these interventions using the formal
system explained above. Our main outcome variable is the mean reduction 
in simulated COVID-19 incidence among group $g$ as a fraction of the no-intervention
cumulative cases, defined at the group level as
\begin{equation}
  \langle \Delta I_g \rangle = \left\langle \frac{I^\mathrm{Tot}_g(\text{no intervention}) -
    I^\mathrm{Tot}_g(\text{intervention } i)}{I^\mathrm{Tot}_g(\text{no
  intervention})} \right\rangle.
  \label{eq:DeltaI_byGroup}
\end{equation}
\noindent
The mean total expected reduction of cumulative cases is 
\begin{equation}
  \langle \Delta I^\mathrm{Tot} \rangle = \left \langle \sum_g \Delta I_g \right \rangle = \sum_g \left \langle \Delta I_g \right \rangle 
  \label{eq:DeltaI}
\end{equation}
\noindent
by linearity of expectations.
In the case of deterministic differential equations, these expectations are just the
values themselves.  We can define A preliminary step is to find parameterizations
that reproduce the observed outbreaks of B.1.1 and Zeta within the Reserve to obtain
a baseline estimate of infection size $I(\text{no intervention})$.  To this end we
hypothesize that re-infections could have driven the observed greater spread of the
Zeta strain, though it could also be due to the holiday season around the end of
2020 and beginning of 2021.

Interventions center around reducing the mobility of Indigenous people between
Dourados city and Reserve; the total change in mobility, denoted $\Delta
\gamma$, is the total reduction of the parameter $\gamma_g$ across all 
groups, i.e., 
\begin{equation}
  \Delta \gamma = \sum_g \Delta \gamma_g.
  \label{eq:DeltaGamma}
\end{equation}
\noindent
Different interventions allocate $\Delta \gamma$ (Equation~\ref{eq:DeltaGamma})
in different ways to hypothetically result in different case reductions for each
group and overall (Equation~\ref{eq:DeltaI_byGroup} and Equation~\ref{eq:DeltaI},
respectfully; see Table~\ref{tab:AnalysisVariables} for a summary of these analysis
variables).
One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
equal. Another named strategy is the \emph{80-20} strategy, where the
reduction in mobility in group $g$ is 80\% of the total, and reduction in
group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$. To create alternative
strategies, both the total reduction and the ratio of group-level reduction to 
total reduction can be changed. In addition to total case reduction, we 
measure relative group-level case reduction, $\Delta I_g$, 
for each considered strategy to understand whether case reduction is equitable
between villages and ethnicities under different strategies. 
We finally evaluate the hypothesis that interventions that may be effective for
strains with no re-infection (like B.1.1) will not be equally effective for
strains with re-infection (like Zeta), and vice-versa. 
% This will highlight the importance of accounting for social and
% epidemiological context, alike, in predicting the effectiveness of mitigation
% strategies for pandemic preparedness.

@@

\vspace{0.5em}
\begin{center}
{[ FIGURE \ref{fig:ByGroupReduction_GlobalStrategy} ABOUT HERE ]} \\
\end{center}

% \vspace{0.5em}
% \begin{center}
% {[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
% \end{center}

\vspace{0.5em}
\begin{center}
{[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
\end{center}

@@

\begin{table}[h]
  \centering
  \caption{\textbf{Independent and dependent analysis variables.} Reducing
  mobility to the city (first row) hypothetically increases the reduction in
cumulative cases in the Reserve (second row).}
  \label{tab:AnalysisVariables}
  \begin{tabular}{rp{3.0in}l} 
    \toprule
    variable & description & values \\
    \midrule 
    $\Delta \gamma$ and $\Delta \gamma_g$ & Total city mobility reduction and 
      group-level mobility reduction & $0.0,0.2,\ldots,0.8$ \\
      $\langle \Delta I^\mathrm{Tot} \rangle$ and $\langle \Delta I^\mathrm{Tot}_g
      \rangle$  & Parameter-set ensemble averages of total and group-level
      cumulative incidence reduction & [0.0, 1.0] \\
    \bottomrule
  \end{tabular} 
\end{table}

@@

% \begin{figure}[H]

% \caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
% strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
% \label{fig:ByGroupReduction_GlobalStrategy}
% \vspace{0.3em}
% {\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}

% }
% \end{figure}

@@

\caption{Disease reduction ($\Delta I$, y-axis) depends on the possibility of
  re-infection and the
distribution strategy to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
\label{fig:ByReinfectionByStrategy}
\vspace{0.3em}
{\centering
  \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByReinfectionByStrategy.pdf}

@@

\caption{Global reduction in travel to and from the city could have
Infection reduction ($\Delta I$, y-axis) for different groups could be
different as global reduction in travel to and from the city decreases
($\Delta \gamma$, x-axis).}
\label{fig:ByGroupReduction_GlobalStrategy}
\vspace{0.3em}
{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ReductionByGroup.pdf}
